Hybrid nonlinear system identification using neural networks

### 2020-11-24 systemID_NNs.tex
ok.

\documentclass{article}
\usepackage[total={6in, 8in}]{geometry}
%% The graphicx and color packages should give you all the color,
%% figure inclusion, and text/figure resizing capabilities you need.
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz,tikzsettings,bm}
\usepackage{almostfull}

@@

%% You should not need more than this for fancy math.
\usepackage{amsmath}   % Extra math commands and environments from the AMS
\usepackage{amssymb}   % Special symbols from the AMS
\usepackage{amsthm}    % Enhanced theorem and proof environments from the AMS
\usepackage{latexsym}  % A few extra LaTeX symbols

\usepackage{lucidbry}
\input{stanacce}
\renewcommand{\baselinestretch}{1.0666}

%% The URL package is handy for typesetting URLs.  It does not define
%% an \email command because so many document styles already do that.
%% So we define one here that uses a typewriter font.
\usepackage{url}
%\DeclareRobustCommand{\email}{\begingroup \urlstyle{tt}\Url}
\providecommand{\email}{}
\renewcommand{\email}[1]{\texttt{#1}}

%% This provides various customized verbatim commands and
%% environments.  You probably don't need it.
\usepackage{fancyvrb}
\DefineShortVerb{\|}
\VerbatimFootnotes
\DefineVerbatimEnvironment{code}{Verbatim}{%
  frame=single,
  framesep=1em,
  xleftmargin=1em,
  xrightmargin=1em,
  samepage=true,
  fontsize=\footnotesize}
\usepackage{upquote}

%% Your document may require a different bibliography style.  I've
%% come to prefer Author (YYYY) styles because it makes it easier for
%% a person who knows something about the literature to understand
%% what is being cited without having to skip to the list of
%% references.
\usepackage[authoryear,round,longnamesfirst]{natbib}

%% You won't normally need this definition in your documents, but it
%% is here so we can typeset the BibTeX logo correctly.
\makeatletter
\@ifundefined{BibTeX}
   {\def\BibTeX{{\rmfamily B\kern-.05em%
    \textsc{i\kern-.025em b}\kern-.08em%
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}}{}

@@

%% Allow more of the pages to be occupied by graphs.  These parameters
%% are described in section c.9.1 of the LaTeXbook.  Your document may
%% not benefit from these parameters (it could make things worse) so
%% you should only change these from the defaults if you need to.
\setcounter{topnumber}{2}              %% 2
\setcounter{bottomnumber}{1}           %% 1
\setcounter{totalnumber}{3}            %% 3
\renewcommand{\topfraction}{0.9}       %% 0.7
\renewcommand{\bottomfraction}{0.9}    %% 0.3
\renewcommand{\textfraction}{0.1}      %% 0.2
\renewcommand{\floatpagefraction}{.7}  %% 0.5
\newlength{\graphwidth}
\setlength{\graphwidth}{0.8\columnwidth}

@@

We augment the grey-box model using
neural networks as follows -
\begin{align} \label{eq:hybrid_cont_time}
\dot x_g &= f_g(x_g, u) + 
f_N(x_g, z, u) \notag \\
y &= h_g(x_g) 
\end{align}
in which $f_N(\cdot)$ is a feedforward neural network
and $z(k) = [\mathbf{y}'_{k-N_p:k-1}, \mathbf{u}'_{k-N_p:k-1}]'$.
This model is converted to discrete-time using the 
Runge-Kutta 4 method, and the vector of past controls
and measurements ($z$) is also viewed as a state. The discrete-time 
model is represented as 
\begin{align*}
  x_g^+ &= f_h(x_g, z, u) \\
  z^+ &= f_z(z, x_g, u)\\
  y &= h_g(x_g) 
\end{align*}
in which, $f_h(\cdot)$ is a obtained by discretization
of \eqref{eq:hybrid_cont_time} using the Runge-Kutta 4 method, 
and $f_z$ is known analytically.

@@

%It is expected that the functions $f_{NN}(\cdot)$ 
%and $h_{NN}(\cdot)$ approximate
%the following differences between the true grey-box state dynamics
%and true measurement function with what is modeled.
%\begin{align*}
%  f_{NN}(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k}) &\approx 
%      Gf_p(x, u) - f_g(Gx, u) \\
%      h_{NN}(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k-1}) &\approx
%      h_p(x) - h_g(Gx) 
%\end{align*}  
%in which $x(k) = \phi(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k-1})$
%is an unknown function of past measurements and inputs.

@@

%\subsection{Linear plant}

%Consider the following plant, model,
%and the expected function to be approximated 
%with a neural network.

%\textbf{Plant --}
%\begin{align*}
%  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
%  \dfrac{dC_B}{dt} &= k_1C_A - k_2C_B + k_3C_C- \dfrac{C_B}{\tau}\\
%  \dfrac{dC_C}{dt} &= k_2C_B - k_3C_C - \dfrac{C_C}{\tau}
%\end{align*}
%These ODEs represent the dynamics of the reaction system
%$A \rightarrow B \rightleftharpoons C$ in a CSTR with constant 
%height and temperature. The control input is $u = C_{Af}$,
%the states are $x = [C_A, C_B, C_C]'$, 
%and the measurements are $y = [C_A, C_B]'$. 
%Assume the following 
%parameters for this model, $k_1 = 1 \ \textnormal{m}^3/\textnormal{min}, 
%\ k_2 = 0.3 \ \textnormal{m}^3/\textnormal{min},
%\ k_3 = 0.2 \ \textnormal{m}^3/\textnormal{min}$, 
%and $\tau = 5 \ \textnormal{min}$.
%The plant model 
%is linear ($dx/dt = Ax + Bu$) and 
%observable with the choices of 
%the parameters and measurements.

%\textbf{Model --}
%\begin{align*}
%  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
%  \dfrac{dC_B}{dt} &= k_1C_A - \dfrac{C_B}{\tau}
%\end{align*}
%i.e, we assume that we don't know about the second reaction while 
%performing grey-box modeling.

%The current plant state $x_p(k)$ can be constructed using the history of 
%past measurements and inputs as 
%\begin{equation*}
%  x_p(k) = A^{N_p-1}_d
%  \mathcal{O}^{\dagger}(\mathbf{y}_{k+1-N_p:k} - 
%                        \mathcal{B}\mathbf{u}_{k+1-N_p:k-1}) + \\ 
%      \mathcal{B}_{nN_p-n+1:nN_p}\mathbf{u}_{k+1-N_p:k-1}
%\end{equation*}

%in which $A_d$ is the plant state transition matrix in discrete-time, 
%$\mathcal{O}$ is the observability matrix, 
%and $\mathcal{B}$ is a combination of the 
%discrete time matrices ($A_d$ and $B_d$) such that it 
%denotes the response of a vector of inputs to states, 
%and $\mathcal{B}_{nN_p-n+1:nN_p}$ denotes the last $n$
%rows of $\mathcal{B}$.

%Since the measurements are a simple linear function 
%of the grey-box and plant states, we use only one 
%neural network to 
%augment the grey-box dynamics. 
%The expected function 
%to be approximated with the network is
%\begin{align*}
%  f_N(\mathbf{y}_{k-N_p:k}, \mathbf{u}_{k+1-N_p:k-1}) \approx 
%  \begin{bmatrix}
%  0 & 0 & 0 \\  
%  0 & -k_2 & k_3 
%  \end{bmatrix}x_p(k)
%\end{align*}

%In this example, the network just approximates a linear function.
%The Figure \ref{fig:validation_linear} 
%below shows validation performances of the grey-box 
%and augmented hybrid models. The structure 
%of the network is restricted to just a matrix and
%the augmented model was trained on 240 samples (4 hours).

%THe Figures \ref{fig:cost_nonlinear_5hours} and 
%\ref{fig:cost_nonlinear_5hours}

%\begin{figure}[!h]
%  \centering
%  \includegraphics[page=2, width=\textwidth]{tworeac_plots_lin.pdf}
%  \caption{Model validation for the linear plant}
%  \label{fig:validation_linear}
%\end{figure}

%We notice that the augmented model accurately predicts the plant 
%behavior. The steady cost profile (Cost = $c_1C_{Af} - c_2C_B$)
%is shown in Figure \ref{fig:cost_linear}.

%\begin{figure}[!h]
%  \centering
%  \includegraphics[page=3, width=\textwidth]{tworeac_plots_lin.pdf}
%  \caption{Steady state cost curve for 
%           the linear plant, grey-box, and hybrid model.}
%  \label{fig:cost_linear}
%\end{figure}

\subsection{Nonlinear reactor example}
We consider a nonlinear reactor with 
two reactions ($A \rightarrow B $ and  $3B \rightleftharpoons C$),
simulated with the following ODEs
\begin{align*}
  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
  \dfrac{dC_B}{dt} &= k_1C_A - 3k_2C^3_B + 3k_3C_C- \dfrac{C_B}{\tau}\\
  \dfrac{dC_C}{dt} &= k_2C^3_B - k_3C_C - \dfrac{C_C}{\tau}
\end{align*}

The grey-box model is kept the same. The parameters 
in the model are modified to 
$k_1 = 1 \ \textnormal{m}^3/\textnormal{min}, 
\ k_2 = 0.01 \ \textnormal{m}^3/\textnormal{min},
\ k_3 = 0.05 \ \textnormal{m}^3/\textnormal{min}$, 
and $\tau = 5 \ \textnormal{min}$.

\newpage
Three models are compared: black-box, residual, and hybrid model with additive 
NN to the grey-box dynamics. The chosen model architectures 
are [9, 16, 16, 2], [9, 16, 16, 2], and [8, 16, 2] respectively.
The Figure \ref{fig:validation_nonlinear} shows model validation on open loop 
data after training with 10 hours of data.

The Figures \ref{fig:cost_nonlinear_5hours} and \ref{fig:cost_nonlinear_10hours}
show the steady state cost curve after training with 5 and 10 hours 
of data respectively. We notice that the hybrid model is more data efficient 
compared to the alternatives.

@@

%We notice that the augmented model accurately predicts the plant 
%behavior. The steady cost profile (Cost = $c_1C_{A0} - c_2C_B$)
%is shown in Figure \ref{fig:cost_nonlinear}. Upon optimization 
%using casadi, the optimum cost and control for the plant are: $-19.62 \$$ 
%and 1.13 $\textnormal{mol}/\textnormal{m}^3$, and for the hybrid 
%grey-box and NN model are: $-19.91 \$$ 
%and 1.11 $\textnormal{mol}/\textnormal{m}^3$.

@@

\begin{figure}[!h]
  \centering
  \includegraphics[page=1, width=\textwidth]{tworeac_distmodels_nonlin.pdf}
  \caption{Steady state grey-box states and disturbance estimates based on 
  output and input disturbance models}
  \label{fig:cost_nonlinear_10hours}
\end{figure}

%\textbf{Need to discuss the following issues:}
%\begin{itemize}
%  \item What are the example problems/applications 
%  where both kind of modeling 
%  strategies make sense, and where should we start.
%  Reproduce results from the Kevrekidis paper with measurement 
%  noise?
%  \item The model training step will be the optimization of 
%  the following multi-step-ahead prediction-error
%  \begin{equation*}
%    \underset{W_i, b_i}{\textnormal{min}} \sum_{k=0}^{t-1} (\hat{y}(k) - y(k))^2
%  \end{equation*}
%  in which ($W_i, b_i$) are the weights of the NN. This optimization 
%  is for one ID experiment, and requires an initial grey-box 
%  state $x_g(0)$ to make the forecast. If we have data 
%  for multiple ID experiments, we'll need initial states for 
%  all the ID experiments. In theory, we can add the initial 
%  states as decision variables in the optimization, but 
%  unfortunately tensorflow 
%  does not has a mechanism to do this.
%\end{itemize}

%\subsection{Adding feedforward disturbance forecasts}
%In this class of problems, we assume that we have a pretty 
%good u-y model from first-principles knowledge, and we'll use 
%neural networks to make feedforward disturbance forecasts
%for use in online MPC. HVAC is a pretty good application.

%\section{Optimization in MPC with neural networks as model}
%This is probably a late fall quarter or a winter break project. \\
%a) Try Casadi/MPC tools. 
%If tanh is the activation function (optimization is smooth),
%this should be fine. \\
%b) Try a global optimization solver. \\
%c) Train another NN with solutions from a global optimization solver.

>>>>

\documentclass{article}
\usepackage[total={4.5in, 8in}]{geometry}
%% The graphicx and color packages should give you all the color, % figure
%inclusion, and text/figure resizing capabilities you need.
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz,tikzsettings,bm}
\usepackage{almostfull}

@@

%% You should not need more than this for fancy math.
\usepackage{amsmath}   % Extra math commands and environments from the AMS
\usepackage{amssymb}   % Special symbols from the AMS
\usepackage{amsthm}    % Enhanced theorem and proof environments from the AMS
\usepackage{latexsym}  % A few extra LaTeX symbols
\usepackage{cleveref}
\usepackage{lucidbry}
\input{stanacce}
\renewcommand{\baselinestretch}{1.0666}

%% The URL package is handy for typesetting URLs.  It does not define % an
%\email command because so many document styles already do that. % So we define
%one here that uses a typewriter font.
\usepackage{url}
%\DeclareRobustCommand{\email}{\begingroup \urlstyle{tt}\Url}
\providecommand{\email}{}
\renewcommand{\email}[1]{\texttt{#1}}

%% This provides various customized verbatim commands and % environments.  You
%probably don't need it.
\usepackage{fancyvrb}
\DefineShortVerb{\|} \VerbatimFootnotes
\DefineVerbatimEnvironment{code}{Verbatim}{%
  frame=single, framesep=1em, xleftmargin=1em, xrightmargin=1em, samepage=true,
  fontsize=\footnotesize}
\usepackage{upquote}

%% Your document may require a different bibliography style.  I've % come to
%prefer Author (YYYY) styles because it makes it easier for % a person who knows
%something about the literature to understand % what is being cited without
%having to skip to the list of % references.
\usepackage[authoryear,round,longnamesfirst]{natbib}

%% You won't normally need this definition in your documents, but it % is here
%so we can typeset the BibTeX logo correctly.
\makeatletter
\@ifundefined{BibTeX} {\def\BibTeX{{\rmfamily B\kern-.05em%
    \textsc{i\kern-.025em b}\kern-.08em%
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}}{}

@@

%% Allow more of the pages to be occupied by graphs.  These parameters % are
%described in section c.9.1 of the LaTeXbook.  Your document may % not benefit
%from these parameters (it could make things worse) so % you should only change
%these from the defaults if you need to.
\setcounter{topnumber}{2}              %% 2
\setcounter{bottomnumber}{1}           %% 1
\setcounter{totalnumber}{3}            %% 3
\renewcommand{\topfraction}{0.9}       %% 0.7
\renewcommand{\bottomfraction}{0.9}    %% 0.3
\renewcommand{\textfraction}{0.1}      %% 0.2
\renewcommand{\floatpagefraction}{.7}  %% 0.5
\newlength{\graphwidth}
\setlength{\graphwidth}{0.8\columnwidth}

@@

We augment the grey-box model using neural networks as follows -
\begin{align} \label{eq:hybrid_cont_time}
\dot x_g &= f_g(x_g, u) + 
f_N(x_g, z, u) \notag \\
y &= h_g(x_g) 
\end{align}
in which $f_N(\cdot)$ is a feedforward neural network and $z(k) =
[\mathbf{y}'_{k-N_p:k-1}, \mathbf{u}'_{k-N_p:k-1}]'$. This model is converted to
discrete-time using the Runge-Kutta 4 method, and the vector of past controls
and measurements ($z$) is also viewed as a state. The discrete-time model is
represented as 
\begin{equation} \label{eq:hybrid_discrete_time}
  \begin{split}
    x_g^+ &= f_h(x_g, z, u)  \\
    z^+ &= f_z(z, x_g, u) \\
    y &= h_g(x_g)       
  \end{split}
\end{equation}
in which, $f_h(\cdot)$ is a obtained by discretization of
\eqref{eq:hybrid_cont_time} using the Runge-Kutta 4 method, and $f_z$ is known
analytically.

@@

@@

\subsection{Nonlinear reactor example}
We consider a nonlinear reactor with two reactions ($A \rightarrow B $ and  $3B
\rightleftharpoons C$), simulated using the following ordinary differential
equations (ODEs)
\begin{align*}
  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
  \dfrac{dC_B}{dt} &= k_1C_A - 3k_2C^3_B + 3k_3C_C- \dfrac{C_B}{\tau}\\
  \dfrac{dC_C}{dt} &= k_2C^3_B - k_3C_C - \dfrac{C_C}{\tau}
\end{align*}

The grey-box model is chosen as follows with an assumption that knowledge of the
second side reaction ($3B \rightleftharpoons C $) is not available
\begin{align*}
  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
  \dfrac{dC_B}{dt} &= k_1C_A - \dfrac{C_B}{\tau}
\end{align*}
The control input is $u = C_{Af}$, the states in the plant model are $x = [C_A,
C_B, C_C]'$, and the measurements are $y = [C_A, C_B]'$. We assume the following
parameters values, $k_1 = 1 \ \textnormal{m}^3/\textnormal{min}, \ k_2 = 0.01 \
\textnormal{m}^3/\textnormal{min}, \ k_3 = 0.05 \
\textnormal{m}^3/\textnormal{min}$, and $\tau = 5 \ \textnormal{min}$.

We examine the performance of the hybrid modeling framework
\eqref{eq:hybrid_discrete_time} in simulation studies. First, we generate
training data by simulating the plant model using a pseudo-random-binary-signal
(PRBS) manipulated input profile. Second, we train a black-box neural network
and the hybrid model \eqref{eq:hybrid_discrete_time} with network architectures
of [9, 16, 2] and [8, 16, 2]. Third, we examine trained model predictions and a
mean squared error (MSE) performance metric on validation data generated using
the plant model. Finally, we analyze the steady state cost curves obtained using
the trained models and their differences with the steady state cost 
curve of the plant model. These
simulations and analysis shed light on the data requirements of the hybrid and black-box
modeling frameworks, and enable an understanding of their performance if used
for real-time economic optimization in chemical plants.

@@

@@

%\begin{figure}[!h] \centering \includegraphics[page=1,
%  width=\textwidth]{tworeac_distmodels_nonlin.pdf} \caption{Steady state
%  grey-box states and disturbance estimates based on output and input
%  disturbance models} \label{fig:cost_nonlinear_10hours} \end{figure}
